---
title: "R Notebook"
output:
  html_document:
    df_print: paged
Writer: Justus Nweze
chunk_output_type: console
Date created: 24 March, 2023
editor_options:
  chunk_output_type: console
---



```{r Load libraries}
# Load the necessary packages
library(readODS)
library(tidyverse)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggthemes)
library(RColorBrewer)
library(vegan)

```





```{r Read the files.}
# 1. Read the files.
metadata <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Metadata")
asv <- read_ods("Metadata_ASV_Taxa.ods", sheet = "ASV")
asv_taxa_filter <- read_ods("Metadata_ASV_Taxa.ods", sheet = "ASV_Taxa_Filter")
```


```{r Normalize}
# Normalize all ASV for each peat soil sample.
# normalize the ASV data by dividing each count by the total count for each sample (column-wise)
asv_norm <- asv
for (col_name in names(asv_norm)[2:28]) {
  asv_norm[[col_name]] <- asv_norm[[col_name]] / sum(asv_norm[[col_name]])
}

# Normalize all ASV for each peat soil sample in asv_taxa_filter using asv sums
asv_taxa_filter_norm <- asv_taxa_filter
for (col_name in names(asv_taxa_filter_norm)[2:28]) {
  asv_taxa_filter_norm[[col_name]] <- asv_taxa_filter_norm[[col_name]] / sum(asv[[col_name]])
}

#In this script, we first create a copy of the asv_taxa_filter data frame called asv_taxa_filter_norm. Then, we use a for loop to iterate through the column names from Floeg093 to Floeg104. For each column, we divide each element in the asv_taxa_filter_norm data frame by the sum of all elements in the same column in the asv data frame. This should now correctly normalize the ASV counts for each peat soil sample in the asv_taxa_filter data frame based on the total counts in the asv data frame and store the result in the asv_taxa_filter_norm data frame.

#The reason for normalizing the original ASV data is to account for differences in sequencing depth between the samples. Different samples may have different total counts, which could affect the comparison of microbial communities. By normalizing the counts, you ensure that all samples are comparable on the same scale.
```



```{r MOB ASV to long, warning=FALSE}
# Convert data in sheet ASV_Taxa_Filter to long, join with Metadata, and summarize.
asv_taxa_filter_long <- asv_taxa_filter_norm %>%
  gather(key = "sample_name", value = "me_count", Floeg094:Floeg104, TF001:TF022) %>%
  left_join(metadata, by = "sample_name") %>%
  mutate(Percentage = me_count * 100)
```





```{r Group by Phylum}
#Group and summarize data by site layer and Phylum

percentage_data_Phylum <- asv_taxa_filter_long %>%
  dplyr::select(sample_name, Phylum, Percentage) %>%
  group_by(sample_name, Phylum) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Phylum) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Phylum, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Phylum$Code_treat <- factor(percentage_data_Phylum$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Phylum.
plot1 <- ggplot(percentage_data_Phylum, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Phylum, fill = Phylum)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 40))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot1)

ggsave("Phylum_Plot.svg", height = 30, width = 100, units = "cm")
```



```{r group by Class}

#Group and summarize data by site layer and Class

percentage_data_Class <- asv_taxa_filter_long %>%
  dplyr::select(sample_name, Class, Percentage) %>%
  group_by(sample_name, Class) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Class) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Class, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Class$Code_treat <- factor(percentage_data_Class$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Class.
plot2 <- ggplot(percentage_data_Class, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Class, fill = Class)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot2)

ggsave("Class_Plot.svg", height = 30, width = 100, units = "cm")


```



```{r group by Order}

# Group and summarize data by site layer and Order

percentage_data_Order <- asv_taxa_filter_long %>%
  dplyr::select(sample_name, Order, Percentage) %>%
  group_by(sample_name, Order) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Order) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Order, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Order$Code_treat <- factor(percentage_data_Order$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Order.
plot3 <- ggplot(percentage_data_Order, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Order, fill = Order)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot3)

ggsave("Order_Plot.svg", height = 30, width = 100, units = "cm")


```



```{r Group by Family}
# Group and summarize data by site layer and family

percentage_data_Family <- asv_taxa_filter_long %>%
  dplyr::select(sample_name, Family, Percentage) %>%
  group_by(sample_name, Family) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Family) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Family, digits = 4)

# Arrange the Code_treat variable according to the desired order
percentage_data_Family$Code_treat <- factor(percentage_data_Family$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Family.
plot4 <- ggplot(percentage_data_Family, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Family, fill = Family)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot4)

ggsave("Family_Plot.svg", height = 30, width = 100, units = "cm")


```



```{r Group by Genus}

# Group and summarize data by site layer and Genus

percentage_data_Genus <- asv_taxa_filter_long %>%
  dplyr::select(sample_name, Genus, Percentage) %>%
  group_by(sample_name, Genus) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat,site_desc, Genus) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Genus, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Genus$Code_treat <- factor(percentage_data_Genus$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Genus.
plot5 <- ggplot(percentage_data_Genus, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Genus, fill = Genus)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot5)

ggsave("Genus_Plot.svg", height = 30, width = 100, units = "cm")




my_palette <- c("#D55E00", "#E69F00", "#009E73", "#F01442", "#000000", "#CC79A7", "#00FFFF", "#999FFF", "#911199", "#444FFC", "#99FF55", "#999999", "#FF11CC")


plot6 <- ggplot(percentage_data_Genus, aes(x = Code_treat, y = mean_Percentage, fill = Genus, color = Genus)) +
  geom_col() +
  scale_fill_manual(values = my_palette[1:13]) +
  scale_color_manual(values = my_palette[1:13]) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50"))

print(plot6)

ggsave("Genus_Plot22.svg", height = 80, width = 100, units = "cm")


# Arrange the Code_treat variable according to the desired order
percentage_data_Genus$Code_treat <- factor(percentage_data_Genus$Code_treat, levels = c("FENpI", "FENpT1","FENpT3", "FENpT4", "SSFT1", "FENpT2", "SSFI", "SSFT2",  "SSFT3",  "SSFT4"))
mycolors <- colorRampPalette(brewer.pal(name = "Paired", n = 12))(13)

plot6_1 <- ggplot(percentage_data_Genus, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(fill = reorder(Genus, mean_Percentage), color = Genus)) +
  scale_fill_manual(values = mycolors) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme_classic() +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50")) #+ facet_wrap(site_desc~Code_treat)

print(plot6_1)





ggsave("Genus_Plot22_2.svg", height = 80, width = 100, units = "cm")








library(kableExtra)
# Group and summarize data by site layer and Genus

percentage_data_Genus <- asv_taxa_filter_long %>%
  dplyr::select(sample_name, Genus, Percentage) %>%
  group_by(sample_name, Genus) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat,site_desc, Genus) %>%
  #dplyr::summarize(mean_Percentage = mean(sum_Percentage), sd_Percentage = sd(sum_Percentage), replicate_number = n()) %>%
  #mutate(se_Percentage = sd_Percentage/sqrt(replicate_number)) %>%
  mutate(Code_treat_Genus = paste(Code_treat, Genus, sep = "_")) %>%
  filter(!is.na(sum_Percentage) & sum_Percentage != 0) %>%  # Remove rows with empty or 0 mean_Percentage
  filter(Genus == "KS41") %>%
  aov(sum_Percentage ~ (Code_treat_Genus),  .) -> # fit an anova model
  aov_t0

summary(aov_t0) 
# print results


TukeyHSD(aov_t0) 


tukey_results <- capture.output(TukeyHSD(aov_t0))
kbl(tukey_results, digits = 4)

```



```{r Indicator Species Analysis}

# Save the reshaped and transposed data frame as ISA_asv_taxa_filter_t
ISA_asv_taxa_filter <- asv_taxa_filter %>%
  select(ASV, 2:28)

# Transpose the reshaped data frame
ISA_asv_taxa_filter_t <- as.data.frame(t(ISA_asv_taxa_filter[-1]))
colnames(ISA_asv_taxa_filter_t) <- ISA_asv_taxa_filter$ASV

# Run the multipatt() function
indval <- multipatt(ISA_asv_taxa_filter_t, metadata$site_desc,
                    control = how(nperm=999))

# Save the summary of indval in indval_summary
summary(indval)

# Get summary information and capture output

output <- capture.output(summary(indval))

# Write output to file
writeLines(output, "Species_Perc_summary.txt")

```





## ** MOB β-diversity indices**

*Distance based matrices such as Bray-Curtis, Unifrac are commonly used and visualized using ordination methods.* 
*Note: The Multidimensionnal Scaling (PCoA) or Principal Coordinates Analysis (PCoA) *

```{r MOB β-diversity indices, cache = T}
# Read in data from file

library("phyloseq")
# load library
library(ggordiplots)

# Load the bacterial abundance table and metadata
ASV_table1 <- read_ods("Metadata_ASV_Taxa.ods", sheet = "ASV_Filter", col_names = TRUE, as_tibble = FALSE, row_names = T, strings_as_factors = T)
Taxa_table1 <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Taxa_Filter", col_names = TRUE, as_tibble = FALSE, row_names = T, strings_as_factors = T)
Peat_metadata1 <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Metadata", col_names = TRUE, as_tibble = FALSE, row_names = T, strings_as_factors = T)

asv_table1 <- as.matrix(ASV_table1)
taxa_table1 <- as.matrix(Taxa_table1)

# Create a phyloseq object
OTU1 <- otu_table(asv_table1, taxa_are_rows = TRUE)
TAX1 <- tax_table(taxa_table1)
sample1 <- sample_data(Peat_metadata1)

carbom1 <- merge_phyloseq(OTU1, TAX1, sample1)
carbom1



# Assume "carbom" is your merged phyloseq object
# Execute prevalence filter, using `prune_taxa()` function
carbom_pruned1 <- prune_taxa(taxa_sums(carbom1) > 0, carbom1)



#*Normalize number of reads in each sample using median sequencing depth.*
total = median(sample_sums(carbom_pruned1))
standf = function(x, t=total) round(t * (x / sum(x)))
carbom_pruned1 = transform_sample_counts(carbom_pruned1, standf)





# Perform NMDS ordination
GP.ord <- ordinate(carbom_pruned1, "NMDS", "bray", k = 3)
#GP.ord <- ordinate(carbom_pruned1, "PCoA", "bray")
# Note the stress value
GP.ord$stress

p11 <- plot_ordination(carbom_pruned1, GP.ord, type="samples", color="Code_treat", shape="site_desc") + 
       geom_point(size=20) + 
       annotate("text", x=-0.5, y=0.4, size = 10, label=paste("stress =", round(GP.ord$stress, 4))) +
       theme(axis.text=element_text(size=40),
             axis.title=element_text(size=40),
             legend.text=element_text(size=40),
             legend.title=element_text(size=40),
             plot.title=element_text(size=40))

print(p11)

ggsave("MOB_anaerobic_Ordination_Plot.svg", height = 60, width = 100, units = "cm")



p111 = plot_ordination(carbom_pruned1, GP.ord, type="sample", color="Code_treat", shape="site_desc") + theme_bw()
p111 + stat_ellipse(geom = "polygon", type="norm", alpha=0.4, aes(fill=Code_treat))




my_palette <- c("#D55E00", "#E69F00", "#009E73", "#F01442", "#000000", "#CC79A7", "#00FFFF", "#999FFF", "#911199", "#444FFC", "#99FF55", "#999999", "#FF11CC")


my_palette <- c("#D55E00", "#009E73", "#F01442", "#FFFF55", "#FF11CC", "#D55E00", "#009E73", "#F01442", "#FFFF55", "#FF11CC")

p12 <- plot_ordination(carbom_pruned1, GP.ord, type = "samples", color = "Code_treat", shape = "site_desc") + 
       geom_point(size = 20) + 
       annotate("text", x = -0.5, y = 0.4, size = 10, label = paste("stress =", round(GP.ord$stress, 4))) +
       theme(axis.text = element_text(size = 40),
             axis.title = element_text(size = 40),
             legend.text = element_text(size = 40),
             legend.title = element_text(size = 40),
             plot.title = element_text(size = 40)) +
       scale_fill_manual(values = my_palette[1:13]) +
       scale_color_manual(values = my_palette[1:13]) + facet_wrap(~site_desc, nrow=1)

print(p12)

ggsave("Anaerobic_MOB_Ordination_Plot.svg", height = 60, width = 100, units = "cm")









library(vegan)
library(factoextra)
library(ggforce)


# Read the metadata and ASV filter data
metadata <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Metadata")
asv_filter <- read_ods("Metadata_ASV_Taxa.ods", sheet = "ASV_Filter")

# Normalize all ASV for each peat soil sample in asv_taxa_filter using asv sums
asv_filter_norm <- asv_filter
for (col_name in names(asv_filter_norm)[2:28]) {
  asv_filter_norm[[col_name]] <- asv_taxa_filter_norm[[col_name]] / sum(asv[[col_name]])
}


# Remove the 'ASV' column from asv_filter to prepare it for merging
asv_filter_long <- asv_filter_norm %>%
  pivot_longer(-ASV, names_to = "sample_name", values_to = "me_count")

# Merge the long-format ASV data with metadata based on 'sample_name'
merged_data <- left_join(asv_filter_long, metadata, by = "sample_name") %>%
  filter(me_count != 0) #%>%
  #filter(site_desc %in% c("SSF")) %>%
  #filter(!grepl("SSFT1|SSFT2|SSFT3", Code_treat))
  #filter(site_desc %in% c("SSF")) %>%
  #filter(!grepl("SSFT1|SSFT2|SSFT3|SSFT4|FENpT1|FENpT2|FENpT3|FENpT4", Code_treat))   # Remove rows with "SSF" in Code_treat

# Create a dissimilarity matrix using Bray-Curtis dissimilarity
diss_matrix <- vegdist(merged_data$me_count, method = "bray")

#=====================================================================================

library(ggforce)







#=====================================================================================
# Group variable for ANOSIM
group_variable <- merged_data$Code_treat

# Perform ANOSIM
anosim_result <- anosim(diss_matrix, group_variable)

# Perform Adonis (PERMANOVA)
adonis_result <- adonis2(diss_matrix ~ Code_treat, data = merged_data)

# Print the ANOSIM result
cat("ANOSIM Result:\n")
print(anosim_result)

# Print the Adonis result
cat("\nAdonis (PERMANOVA) Result:\n")
print(adonis_result)





nmds <- metaMDS(diss_matrix)

# Plot the NMDS results
plot(nmds, type = "t")  # "n" to create an empty plot first
points(nmds, col = merged_data$site_desc)  # Color by site description

# Add labels to the points if needed
text(nmds, labels = rownames(merged_data), cex = 0.8)

```








```{r Read the files2.}
# Analyze all microbial community

# Read the files.
metadata <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Metadata")

asv <- read_ods("Metadata_ASV_Taxa.ods", sheet = "ASV")

Taxa <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Taxa")

```




```{r Normalize2}
# Normalize all ASV for each peat soil sample.
# normalize the ASV data by dividing each count by the total count for each sample (column-wise)
asv_norm <- asv
for (col_name in names(asv_norm)[2:28]) {
  asv_norm[[col_name]] <- asv_norm[[col_name]] / sum(asv_norm[[col_name]])
}


asv_taxa_norm <- merge(asv_norm, Taxa, by = "ASV")

```



```{r ASV to long2, warning=FALSE}
# Convert data in sheet ASV to long, join with Metadata, and summarize.
# Assuming your data frames are named metadata, asv_norm

# Convert data in sheet ASV to long format using pivot_longer function
asv_taxa_norm_long <- asv_taxa_norm %>%
  pivot_longer(cols = Floeg094:TF022, names_to = "sample_name", values_to = "Total_count")

# Join it with data in sheet Metadata using sample_name
asv_taxa_norm_long_joined <- asv_taxa_norm_long %>%
  left_join(metadata, by = "sample_name") %>%
  mutate(Percentage = Total_count * 100)

```






```{r All microbial Group by Phylum}
#Group and summarize data by site layer and Phylum

percentage_data_All_Phylum <- asv_taxa_norm_long_joined%>%
  dplyr::select(sample_name, Phylum, Percentage) %>%
  group_by(sample_name, Phylum) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Phylum) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_All_Phylum, digits = 4)


  
# Select the top 10 Phyla based on mean relative abundance across all samples
top_phyla <- percentage_data_All_Phylum %>%
  group_by(Phylum) %>%
  filter(mean_Percentage > 1)

# Filter the data to keep only the top Phyla
percentage_data_All_Phylum_filtered <- percentage_data_All_Phylum %>%
  filter(Phylum %in% top_phyla$Phylum)

# Arrange the sample_name variable according to the desired order
percentage_data_All_Phylum_filtered$Code_treat <- factor(percentage_data_All_Phylum_filtered$Code_treat, levels = c("FENpI", "FENpT1", "FENpT2", "FENpT3","FENpT4","SSFI", "SSFT1", "SSFT2","SSFT3", "SSFT4"))

my_palette <- c(
  "#FF00FF", "#FF0000", "#000000", "#008B8B", "#FFC0CB", "#00008B", "#F0F8FF", "#8B008B",
  "#DC143C", "#C0C0C0", "#00FFFF", "#556B2F", "#C71585", "#FFFF00", "#0000FF"
)  


# Create the plot by Phylum.
plot11_1 <- ggplot(percentage_data_All_Phylum_filtered, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Phylum, fill = Phylum)) +
 scale_fill_manual(values = my_palette[1:21]) +
  scale_color_manual(values = my_palette[1:21]) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50"))

print(plot11_1)

ggsave("All_Top_Phylum_Plot.svg", height = 50, width = 100, units = "cm")

mycolors <- colorRampPalette(brewer.pal(name = "Paired", n = 12))(24)

# Create the plot by Phylum.
plot11_2 <- ggplot(percentage_data_All_Phylum_filtered, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(fill = reorder(Phylum, mean_Percentage), color = Phylum)) +
 scale_fill_manual(values = mycolors) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50"))

print(plot11_2)

ggsave("All_Top_Phylum_Plot_22.svg", height = 50, width = 100, units = "cm")
```




```{r All microbial Group by Genus}
#Group and summarize data by site layer and Genus

percentage_data_All_Genus <- asv_taxa_norm_long_joined%>%
  dplyr::select(sample_name, Genus, Percentage) %>%
  group_by(sample_name, Genus) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Genus) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
kbl(percentage_data_All_Genus, digits = 4)



top_genera <- percentage_data_All_Genus %>%
  group_by(Genus) %>%
  filter(mean_Percentage > 1)

# Filter the data to keep only the top Phyla
percentage_data_All_Genus_filtered <- percentage_data_All_Genus %>%
  filter(Genus %in% top_genera$Genus)

# Arrange the sample_name variable according to the desired order
percentage_data_All_Genus_filtered$Code_treat <- factor(percentage_data_All_Genus_filtered$Code_treat, levels = c("FENpI", "FENpT1", "FENpT2", "FENpT3","FENpT4","SSFI", "SSFT1", "SSFT2","SSFT3", "SSFT4"))

# Create the plot by Genus.


my_palette1 <- c(
  
"#556B2F", "#8B0000", "#000000", "#008B8B", "#C71585", "#666600", "#00008B", "#4B0082", "#8B008B",
"#2E8B57", "#DC143C", "#696969", "#5F9EA0", "#FF1493", "#FFFF00", "#0000FF", "#8A2BE2", "#FF00FF",
"#98FB98", "#FF6347", "#C0C0C0", "#00FFFF", "#FF69B4", "#FFFF99", "#1E90FF", "#6495ED", "#EE82EE",
"#7CFC00", "#FF0000", "#ffffff", "#E0FFFF", "#FFC0CB", "#FFFFCC", "#ADD8E6", "#F0F8FF", "#D8BFD8")


my_palette <- c(
  "#556B2F", "#FF0000", "#000000", "#008B8B", "#FFC0CB", "#666600", "#00008B", "#F0F8FF", "#8B008B",
  "#7CFC00", "#DC143C", "#C0C0C0", "#00FFFF", "#C71585", "#FFFF00", "#ADD8E6", "#8A2BE2", "#FF00FF",
  "#ffffff", "#FFFFCC", "#0000FF"
)  


plot12 <- ggplot(percentage_data_All_Genus_filtered, aes(x = Code_treat, y = mean_Percentage, fill = reorder(Genus, mean_Percentage), color = Genus)) +
  geom_col() +
 scale_fill_manual(values = my_palette[1:21]) +
  scale_color_manual(values = my_palette[1:21]) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50"))

print(plot12)

ggsave("All_Top_Genera_Plot1.svg", height = 30, width = 100, units = "cm")


mycolors <- colorRampPalette(brewer.pal(name = "Paired", n = 12))(55)

plot12_1 <- ggplot(percentage_data_All_Genus_filtered, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(fill = reorder(Genus, mean_Percentage), color = Genus)) +
 scale_fill_manual(values = mycolors) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50"))

print(plot12_1)


ggsave("All_Top_Genera_Plot22.svg", height = 50, width = 100, units = "cm")

```



## **β-diversity indices**

*Distance based matrices such as Bray-Curtis, Unifrac are commonly used and visualized using ordination methods.* 
*Note: The Multidimensionnal Scaling (PCoA) or Principal Coordinates Analysis (PCoA) *
```{r All microbial β-diversity indices, cache = T}
# Read in data from file
# Load the bacterial abundance table and metadata
ASV_table <- read_ods("Metadata_ASV_Taxa.ods", sheet = "ASV", col_names = TRUE, row_names = T, strings_as_factors = T)
Taxa_table <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Taxa", col_names = TRUE, row_names = T, strings_as_factors = T)
Peat_metadata <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Metadata", col_names = TRUE, row_names = T, strings_as_factors = T)

asv_table <- as.matrix(ASV_table)
taxa_table <- as.matrix(Taxa_table)

# Create a phyloseq object
OTU <- otu_table(asv_table, taxa_are_rows = TRUE)
TAX <- tax_table(taxa_table)
sample <- sample_data(Peat_metadata)

carbom <- merge_phyloseq(OTU, TAX, sample)
carbom

# Assume "carbom" is your merged phyloseq object
# Execute prevalence filter, using `prune_taxa()` function
carbom_pruned <- prune_taxa(taxa_sums(carbom) > 0, carbom)


#*Normalize number of reads in each sample using median sequencing depth.*
total = median(sample_sums(carbom_pruned))
standf = function(x, t=total) round(t * (x / sum(x)))
carbom_pruned = transform_sample_counts(carbom_pruned, standf)


GP.ord <- ordinate(carbom_pruned, "NMDS", "bray")
p1 = plot_ordination(carbom_pruned, GP.ord, type="samples", color="Code_treat", shape="site_desc") + geom_point(size=50) + 
       annotate("text", x=-0.4, y=0.57, size = 20, label=paste("stress =", round(GP.ord$stress, 4))) +
       theme(axis.text=element_text(size=40),
             axis.title=element_text(size=40),
             legend.text=element_text(size=40),
             legend.title=element_text(size=40),
             plot.title=element_text(size=40))

print(p1)
ggsave("All_anaerobic_Microbial3_Ordination_Plot.svg", height = 50, width = 100, units = "cm")



my_palette <- c("#D55E00", "#009E73", "#F01442", "#FFFF55", "#FF11CC", "#D55E00", "#009E73", "#F01442", "#FFFF55", "#FF11CC")

p120 <- plot_ordination(carbom_pruned, GP.ord, type = "samples", color = "Code_treat", shape = "site_desc") + 
       geom_point(size = 50) + 
       annotate("text", x = 1, y = 1, size = 10, label = paste("stress =", round(GP.ord$stress, 4))) +
       theme(axis.text = element_text(size = 40),
             axis.title = element_text(size = 40),
             legend.text = element_text(size = 40),
             legend.title = element_text(size = 40),
             plot.title = element_text(size = 40)) +
       scale_fill_manual(values = my_palette[1:13]) +
       scale_color_manual(values = my_palette[1:13])

print(p120)

ggsave("All_anaerobic_Microbial_Ordination2_Plot.svg", height = 80, width = 100, units = "cm")
```































# For Archaea


```{r Read the files.}
# 1. Read the files.
metadata <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Metadata")
asv <- read_ods("Metadata_ASV_Taxa.ods", sheet = "ASV")
Archaea <- read_ods("Metadata_ASV_Taxa.ods", sheet = "Archaea")
```


```{r Normalize}
# Normalize all ASV for each peat soil sample.
# normalize the ASV data by dividing each count by the total count for each sample (column-wise)
asv_norm <- asv
for (col_name in names(asv_norm)[2:28]) {
  asv_norm[[col_name]] <- asv_norm[[col_name]] / sum(asv_norm[[col_name]])
}

# Normalize all ASV for each peat soil sample in Archaea using asv sums
Archaea_norm <- Archaea
for (col_name in names(Archaea_norm)[2:28]) {
  Archaea_norm[[col_name]] <- Archaea_norm[[col_name]] / sum(asv[[col_name]])
}

#In this script, we first create a copy of the Archaea data frame called Archaea_norm. Then, we use a for loop to iterate through the column names from Floeg093 to Floeg104. For each column, we divide each element in the Archaea_norm data frame by the sum of all elements in the same column in the asv data frame. This should now correctly normalize the ASV counts for each peat soil sample in the Archaea data frame based on the total counts in the asv data frame and store the result in the Archaea_norm data frame.

#The reason for normalizing the original ASV data is to account for differences in sequencing depth between the samples. Different samples may have different total counts, which could affect the comparison of microbial communities. By normalizing the counts, you ensure that all samples are comparable on the same scale.
```



```{r MOB ASV to long, warning=FALSE}
# Convert data in sheet Archaea to long, join with Metadata, and summarize.
Archaea_long <- Archaea_norm %>%
  gather(key = "sample_name", value = "me_count", Floeg094:Floeg104, TF001:TF022) %>%
  left_join(metadata, by = "sample_name") %>%
  mutate(Percentage = me_count * 100)
```





```{r Group by Phylum}
#Group and summarize data by site layer and Phylum

percentage_data_Phylum <- Archaea_long %>%
  dplyr::select(sample_name, Phylum, Percentage) %>%
  group_by(sample_name, Phylum) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Phylum) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Phylum, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Phylum$Code_treat <- factor(percentage_data_Phylum$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Phylum.
plot1 <- ggplot(percentage_data_Phylum, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Phylum, fill = Phylum)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 40))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot1)

ggsave("Archaea_Phylum_Plot.svg", height = 30, width = 100, units = "cm")
```



```{r group by Class}

#Group and summarize data by site layer and Class

percentage_data_Class <- Archaea_long %>%
  dplyr::select(sample_name, Class, Percentage) %>%
  group_by(sample_name, Class) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Class) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Class, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Class$Code_treat <- factor(percentage_data_Class$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Class.
plot2 <- ggplot(percentage_data_Class, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Class, fill = Class)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot2)

ggsave("Archaea_Class_Plot.svg", height = 30, width = 100, units = "cm")


```



```{r group by Order}

# Group and summarize data by site layer and Order

percentage_data_Order <- Archaea_long %>%
  dplyr::select(sample_name, Order, Percentage) %>%
  group_by(sample_name, Order) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Order) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Order, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Order$Code_treat <- factor(percentage_data_Order$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Order.
plot3 <- ggplot(percentage_data_Order, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Order, fill = Order)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot3)

ggsave("Archaea_Order_Plot.svg", height = 30, width = 100, units = "cm")


```



```{r Group by Family}
# Group and summarize data by site layer and family

percentage_data_Family <- Archaea_long %>%
  dplyr::select(sample_name, Family, Percentage) %>%
  group_by(sample_name, Family) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat, Family) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Family, digits = 4)

# Arrange the Code_treat variable according to the desired order
percentage_data_Family$Code_treat <- factor(percentage_data_Family$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Family.
plot4 <- ggplot(percentage_data_Family, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Family, fill = Family)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot4)

ggsave("Archaea_Family_Plot.svg", height = 30, width = 100, units = "cm")


```



```{r Group by Genus}

# Group and summarize data by site layer and Genus

percentage_data_Genus <- Archaea_long %>%
  dplyr::select(sample_name, Genus, Percentage) %>%
  group_by(sample_name, Genus) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat,site_desc, Genus) %>%
  dplyr::summarize(mean_Percentage = mean(sum_Percentage), .groups = 'drop')    
  kbl(percentage_data_Genus, digits = 4)


# Arrange the Code_treat variable according to the desired order
percentage_data_Genus$Code_treat <- factor(percentage_data_Genus$Code_treat, levels = c("FENpI", "SSFI", "FENpT1", "SSFT1", "FENpT2", "SSFT2", "FENpT3", "SSFT3", "FENpT4", "SSFT4"))

# Create the plot by Genus.
plot5 <- ggplot(percentage_data_Genus, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(color = Genus, fill = Genus)) +
  ylab("Relative Abundance (%)") +
  theme(text = element_text(size = 40)) +
  theme(axis.text = element_text(size = 30))
# with position = "fill", the bars are normalized so that they represent relative abundance (percentages) within each environment:
print(plot5)

ggsave("Archaea_Genus_Plot.svg", height = 30, width = 100, units = "cm")




my_palette <- c("#D55E00", "#E69F00", "#009E73", "#F01442", "#000000", "#CC79A7", "#00FFFF", "#999FFF", "#911199", "#444FFC", "#99FF55", "#999999", "#FF11CC")

my_palette1 <- c(
  
"#556B2F", "#8B0000", "#000000", "#008B8B", "#C71585", "#666600", "#00008B", "#4B0082", "#8B008B",
"#2E8B57", "#DC143C", "#696969", "#5F9EA0", "#FF1493", "#FFFF00", "#0000FF", "#8A2BE2", "#FF00FF",
"#98FB98", "#FF6347", "#C0C0C0", "#00FFFF", "#FF69B4", "#FFFF99", "#1E90FF", "#6495ED", "#EE82EE",
"#7CFC00", "#FF0000", "#ffffff", "#E0FFFF", "#FFC0CB", "#FFFFCC", "#ADD8E6", "#F0F8FF", "#D8BFD8")


plot6 <- ggplot(percentage_data_Genus, aes(x = Code_treat, y = mean_Percentage, fill = Genus, color = Genus)) +
  geom_col() +
  scale_fill_manual(values = my_palette[1:24]) +
  scale_color_manual(values = my_palette[1:24]) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50"))

print(plot6)

ggsave("Archaea_Genus_Plot22.svg", height = 80, width = 100, units = "cm")


# Arrange the Code_treat variable according to the desired order
percentage_data_Genus$Code_treat <- factor(percentage_data_Genus$Code_treat, levels = c("FENpI", "FENpT1","FENpT3", "FENpT4", "SSFT1", "FENpT2", "SSFI", "SSFT2",  "SSFT3",  "SSFT4"))
mycolors <- colorRampPalette(brewer.pal(name = "Paired", n = 12))(13)

plot6_1 <- ggplot(percentage_data_Genus, aes(x = Code_treat, y = mean_Percentage)) +
  geom_col(aes(fill = reorder(Genus, mean_Percentage), color = Genus)) +
  scale_fill_manual(values = mycolors) +
  ylab("Relative Abundance (%)") +
  theme(axis.text = element_text(size = 40)) +
  theme_classic() +
  theme(panel.border = element_blank(),
        text = element_text(size = 40),
        panel.background = element_rect(fill = "white", colour = "grey50")) #+ facet_wrap(site_desc~Code_treat)

print(plot6_1)





ggsave("Genus_Plot22_2.svg", height = 80, width = 100, units = "cm")








library(kableExtra)
# Group and summarize data by site layer and Genus

percentage_data_Genus <- Archaea_long %>%
  dplyr::select(sample_name, Genus, Percentage) %>%
  group_by(sample_name, Genus) %>%
  dplyr::summarize(sum_Percentage = sum(Percentage), .groups = 'drop') %>% 
  left_join(metadata, by = "sample_name") %>%                             
  group_by(Code_treat,site_desc, Genus) %>%
  #dplyr::summarize(mean_Percentage = mean(sum_Percentage), sd_Percentage = sd(sum_Percentage), replicate_number = n()) %>%
  #mutate(se_Percentage = sd_Percentage/sqrt(replicate_number)) %>%
  mutate(Code_treat_Genus = paste(Code_treat, Genus, sep = "_")) %>%
  filter(!is.na(sum_Percentage) & sum_Percentage != 0) %>%  # Remove rows with empty or 0 mean_Percentage
  filter(Genus == "KS41") %>%
  aov(sum_Percentage ~ (Code_treat_Genus),  .) -> # fit an anova model
  aov_t0

summary(aov_t0) 
# print results


TukeyHSD(aov_t0) 


tukey_results <- capture.output(TukeyHSD(aov_t0))
kbl(tukey_results, digits = 4)

```
